\documentclass{notes}
\begin{document}
\vspace*{\fill}
\begin{center}
{\centering\huge{MECH 423 Proposal}}\\
\vspace*{2em}
{\large Jasper Chan: 37467164 }\\
{\large Jacky Chang: 19835164 }
\end{center}
\vspace*{\fill}
\newpage
\section{Objectives}

The objective of this project is to design and create a simple two player game similar to LightPong\footnotemark.
\footnotetext{
    \url{https://www.kickstarter.com/projects/aaqib/light-pong-game}
}
The system will consist of two handles/controllers on each end of a flexible tube that houses an LED strip.
The MSP430 will be used to control the LED's and process the signals from each controller.

\section{Rationale}

It's interesting to reverse engineer a product and see what type of difficulties the original developers may have experienced along the way.
The product is still currently in it's crowdfunding phase, so it's not in full development, and it would be nice to be able to have one to play with.

\section{List of Functional Requirements}

\begin{table}[H]
    \centering
    \begin{tabularx}{\textwidth}
        {
            >{\hsize=.6\hsize\raggedright}X
            >{\hsize=.1\hsize\raggedright}X
            >{\hsize=.3\hsize\raggedright\arraybackslash}X
        }
        Task                                                                    & Effort            & Person Responsible \\\toprule
        Power + charging PCB capable of delivering enough current for LED strip & \SI{20}{\percent} & Jasper             \\\midrule
        Microcontroller PCB to drive LED strip and read button inputs           & \SI{20}{\percent} & Jasper             \\\midrule
        Mechanical housing for parts                                            & \SI{20}{\percent} & Jacky              \\\midrule
        Addressable LED Strip                                                   & \SI{05}{\percent} & Jacky              \\ \midrule
        Microcontroller code                                                    & \SI{35}{\percent} & Jasper and Jacky   \\\bottomrule
    \end{tabularx}
    \caption{Functional Requirements and their efforts}
    \label{tab:funcreq}
\end{table}

\section{Functional Requirement 1: Addressable LED strip}

\subsection{Approach and Design}

The most common addressable LED strips on the market are all based on the WS2812B (commonly referred to as Neopixels).
The general rule of thumb is that each LED may consume up to \SI{60}{\milli\ampere} of current, although this rarely happens when displaying patterns on the strip since not all LEDs will be on/at full power.
For our application, Adafruit sells\footnotemark\ a \SI{1}{\meter} long strip with 30 LEDs.
\footnotetext{\url{https://www.adafruit.com/product/1376}}
This provides an acceptable balance between LED density and current consumption.

These LEDs are typically rated for \SI{5}{\volt} power and logic, however they can commonly be driven at lower voltages.
In order to determine the power requirements for the power PCB, experimentation will have to be done.

\subsection{Inputs and Outputs}

\subsubsection{Inputs}

\begin{enumerate}
    \item Power
    \begin{enumerate}
        \item Voltage: nominally \SI{5}{\volt}, actual TBD
        \item Current: Max \SI{60}{\milli\ampere} per LED
    \end{enumerate}
    \item Data: the WS2812B data format is described in the datasheet\footnotemark
    \footnotetext{\url{https://cdn-shop.adafruit.com/datasheets/WS2812.pdf}}
    \item Wiring: Wires between the handles will need to be run along the LED strip
\end{enumerate}

\subsubsection{Outputs}

\begin{enumerate}
    \item Light: the LED strip will emit light
    \item Power: the power wiring inside the LED strip may be used to power the microcontroller PCB
    \item Wiring: Wires between the handles will need to be run along the LED strip
\end{enumerate}

\subsection{Parameters}

\begin{enumerate}
    \item Strip length
    \begin{enumerate}
        \item Adafruit sells strips in \SI{1}{\meter} segments, but they can be cut in between each LED
        \item A \SI{1}{\meter} strip will be used for convenience
    \end{enumerate}
    \item LED count
    \begin{enumerate}
        \item Adafruit sells strips at various LED densities
        \item Higher densities will increase power and memory requirements on the MCU
    \end{enumerate}
\end{enumerate}

\subsection{Development Plan}

For simplicity, a \SI{1}{\meter} strip with 30 LEDs will be used.

A \SI{3.3}{\volt} logic level microcontroller will be used to determine if the strip can be controlled without the use of a level shifter.

A bench power supply will be used to determine if a \SI{5}{\volt} boost circuit will be needed to power the LEDs.

\subsection{Test Plan}
\label{sec:fr1testplan}

\subsubsection{Logic level requirements}
\begin{enumerate}
    \item Power the LED strip with \SI{5}{\volt} from a bench power supply
    \item Connect a \SI{3.3}{\volt} logic level MCU to the data pin of the strip
    \item Attempt to display a test pattern, if this works, a level shifter is not required, if this does not work, redo the test with a level shifter.
\end{enumerate}

\subsubsection{Brightness requirements}
\begin{enumerate}
    \item Power the LED strip with \SI{5}{\volt} from a bench power supply
    \item Connect an MCU to the data pin of the strip
    \item Set all LEDs to full brightness white (RGB 255, 255, 255)
    \item Incrementally turn down the brightness until it is at a reasonable level
\end{enumerate}

\subsubsection{Voltage requirements}
\begin{enumerate}
    \item Power the LED strip with \SI{3.2}{\volt} from a bench power supply
    \item Connect an MCU to the data pin of the strip
    \item Set all LEDs to a reasonable brightness of white
    \item If the LEDs turn on correctly, note the current consumption, if not, repeat the test at a higher voltage until the LED strip reliably turns on.
    \begin{itemize}
        \item This will determine if we need a boost circuit or if the LEDs can be directly powered from a LiPo battery.
    \end{itemize}
\end{enumerate}

\subsection{Integration}
\begin{enumerate}
    \item Power the LED strip with the determined acceptable voltage level
    \item Connect MCU (with level shifter if required) to the data pin of the strip
    \item Display a test pattern on the strip
\end{enumerate}

\section{Functional Requirement 2: Power + Charging PCB}
\label{sec:fr2}

\subsection{Approach and Design}

Because there is limited space to store a power source within the handles of the device, a custom PCB will have to be made to hold the battery, and associated charging, protection and boost circuitry as required.

Because the handle will be cylindrical, an 18650 battery will be used because it will be the most convenient form factor while providing high power and energy density.
For safety reasons, a spring-loaded battery holder will be used to avoid soldering or spot welding connections.

\subsection{Inputs and Outputs}

\subsubsection{Inputs}

\begin{enumerate}
    \item Charging power: Battery will be charged with a USB connection
    \item Button: Power PCB will be housed in one of the device handles, and will need to have a connector for the handle button
    \item Button LED signal: Power PCB will need a connection to receive the signal to turn the button's LED on or off.
    \item Power Switch: The power PCB will need a power switch to disable the power output.
\end{enumerate}

\subsubsection{Outputs}

\begin{enumerate}
    \item LED strip power: Power PCB will need to provide power to the LED strip
    \item Button signal: Power PCB will need to have a connection to deliver the button signal to the MCU
    \item Button LED power: Power PCB will need to light up the button's LED when indicated by the MCU
\end{enumerate}

\subsection{Parameters}

\begin{enumerate}
    \item LED strip voltage: Determined in \autoref{sec:fr1testplan}
    \item Total sustained current: Determined in \autoref{sec:fr1testplan}
    \item Battery choice: Must be capable of providing enough power
    \begin{itemize}
        \item Battery may or may not come with protection circuitry built in.
        Requirement for protection circuitry on the PCB is only required for unprotected cells.
        PCB silkscreen will clearly indicate whether protected or unprotected cells should be used.
    \end{itemize}
    \item Charging current: for safety reasons this will be a conservative number (max 1C)
    \item PCB dimensions: PCB must be big enough to hold all the components
\end{enumerate}

\subsection{Development Plan}

\begin{enumerate}
    \item Design schematic and layout of PCB
    \item Send design to PCB house for manufacturing, buy components
    \item Assemble PCB
\end{enumerate}

\subsection{Test Plan}

\subsubsection{Charging}

\begin{enumerate}
    \item Place PCB and USB battery bank inside a fireproof container
    \item Connect USB battery bank to PCB charging input
    \item Wait for charging to complete.
    \item Remove LiPo cell from PCB, ensure that it has the correct voltage for a charged cell (approximately \SI{4.2}{\volt})
\end{enumerate}

\subsubsection{Power Output}

\begin{enumerate}
    \item Place PCB inside a fireproof container
    \item Connect PCB power output to a load
        \begin{itemize}
            \item If the lab contains a programmable load, this is preferred, otherwise the load can just be the LED strip + MCU
        \end{itemize}
    \item Enable PCB power output and load
    \item Ensure power output is stable
    \item Keep test running, ensure power output stops at or before designed battery cutoff voltage is reached.
\end{enumerate}

\subsubsection{Button LED}

\begin{enumerate}
    \item Connect button to PCB
    \item Connect MCU GPIO to PCB
    \item Enable PCB power output
    \item Toggle Button LED GPIO, ensure LED toggles.
\end{enumerate}

\subsubsection{Button Input}

\begin{enumerate}
    \item Connect button to PCB
    \item Connect MCU GPIO to PCB
    \item Enable PCB power output
    \item Press and release button, ensure MCU can read button input.
\end{enumerate}

\section{Functional Requirement 3: Microcontroller PCB}

\subsection{Approach and Design}

As required by the class, an MSP430 will be used.
For simplicity, the same MSP430FR5739IDAR\footnotemark as used in Lab 3 will be used, and the design will be derived from the Lab 3 design.
\footnotetext{
    \url{https://www.digikey.ca/en/products/detail/texas-instruments/MSP430FR5739IDAR/2696032}
}

\subsection{Inputs and Outputs}

\subsubsection{Inputs}

\begin{enumerate}
    \item Power: Power will be provided from the PCB board, either via the LED strip, or via a separate wire. Power may have to be stepped down to safely power the MCU
    \item Button x2: MCU PCB will be housed in one of the device handles, and will need to have a connector for the handle button, as well as for the button on the other handle
\end{enumerate}

\subsubsection{Outputs}

\begin{enumerate}
    \item LED strip data: MCU PCB will need to control the LED strip
    \item Button LED power x2: MCU PCB will need to light up the button's LED when indicated by the MCU
\end{enumerate}

\subsection{Parameters}

\begin{enumerate}
    \item Crystal choice: Determine frequency and precision of MCLK source
\end{enumerate}

\subsection{Development Plan}

\begin{enumerate}
    \item Design schematic and layout of PCB
    \item Send design to PCB house for manufacturing, buy components
    \item Assemble PCB
\end{enumerate}

\subsection{Test Plan}

\begin{enumerate}
    \item Power up PCB with bench power supply (use low current limit to minimize risk of shorts)
    \item Check that MCU can be programmed
    \item Program MCU with test firmware
    \item Confirm functionality of all required IO pins with a test program
\end{enumerate}

\section{Functional Requirement 4: Mechanical Housing}

\subsection{Approach and Design}

A custom mechanical housing will have to be designed to house the other components.
To avoid spending too much time or money machining, 3D printing will be primarily used to fabricate mechanical parts.
To make debugging of other components easier, the housing will be designed to be disassembled easily.

\subsection{Inputs and Outputs}

\subsubsection{Inputs}

\begin{enumerate}
    \item Other components: all other components will either be fixed to or contained within the mechanical housing
\end{enumerate}

\subsubsection{Outputs}

\begin{enumerate}
    \item Mechanical stability: parts must be held securely
\end{enumerate}

\subsection{Parameters}

\begin{enumerate}
    \item Handle diameter: diameter must be small enough to be comfortable, but large enough contain the PCBs
    \item Handle length: length must be short enough to be comfortable, but long enough to contain the PCBs
\end{enumerate}

\subsection{Development Plan}

\begin{enumerate}
    \item Once PCBs have been designed, PCB mechanical models will be passed on to the mechanical housing designer
    \item Mechanical housing will be designed around the PCB models
\end{enumerate}

\subsection{Test Plan}

\begin{enumerate}
    \item PCBs (or printed models of the PCBs) will be assembled into the mechanical housings to test fit
\end{enumerate}

\section{Functional Requirement 5: Microcontroller Code}

\subsection{Approach and Design}

The code for the MCU will need to be written to implement the game behavior.
Once the schematic for the MCU PCB is finalized, code can be written to run on it.

\subsection{Inputs and Outputs}

\subsubsection{Inputs}

\begin{enumerate}
    \item Buttons: The code must read the button inputs at the appropriate time
\end{enumerate}

\subsubsection{Outputs}

\begin{enumerate}
    \item Button LEDs: The code must drive the button LEDs at the appropriate time
    \item LED strip: The code must drive the Neopixel LED strip at the appropriate time with the correct pattern
\end{enumerate}

\subsection{Parameters}

\begin{enumerate}
    \item Game speed: The movement speed of the "ball" will control how hard the game is to play
\end{enumerate}

\subsection{Development Plan}

\begin{enumerate}
    \item Once the schematic is finalized, code can be written for the PCB
\end{enumerate}

\subsection{Test Plan}

\begin{enumerate}
    \item Flash the code on the MCU PCB (or another MSP430 board if the MCU PCB has not been completed)
    \item Hook up the buttons and LED strip and test that the code behaves as expected
\end{enumerate}

\section{Most Critical Module}

The most critical module is the power PCB described in \autoref{sec:fr2}.
The power PCB needs to provide stable power to all the other components.
The power PCB will also be the larger of the two PCBs, and will be the limiting factor in how small the handles can be.

\section{Risks and Countermeasures}

\subsection{Parts Shortage}

We are in the middle of a parts and shipping shortage at time of writing.

To mitigate potential shipping delays, designs will be completed as soon as possible so that parts can be ordered as soon as possible.
To mitigate parts potentially being available, a BOM will be kept up to date as the design progresses, and will be periodically checked with DigiKey to ensure that they are still available.
Where possible, PCBs will be designed to support alternative parts.
The most critical part will be the MSP430FR5739IDAR, which at time of writing is still available.
If it is no longer available at time of purchasing, the chip can be salvaged from the PCB from Lab 3.

\subsection{Limited Power}

The battery or circuitry in the power PCB may not deliver enough power to the rest of the system when assembled.

To mitigate limited power output, the USB charging connector and an auxiliary barrel connector will also be routed to the main power output so that the system can be externally powered.

\section{Learning Objectives}

We hope to further develop both electrical and mechanical CAD skills, as well as programming skills.

\end{document}
